Forecasting the 2024 U.S. Presidential Election: A Poll-of-Polls Approach for Predicting the Outcome

### 2024-11-04 paper/paper.qmd
result change

- **Average Support**: Kamala Harris has a slight lead over Donald Trump, with an average predicted national support of 47.39% compared to Trump’s 45.65%. This narrow margin points to a competitive race, with neither candidate establishing a decisive lead across the polls.

- **Variability in Support**: The standard deviation values indicate greater variability in Trump’s support (5.33%) compared to Harris’s (4.38%). This suggests that Trump’s polling performance fluctuates more widely, which could be due to varying levels of regional support or shifts in public opinion over time.

- **Support Range**: Both candidates display a considerable range in support across polls, with Trump’s polling support varying between 21.0% and 70.0%, and Harris’s between 25.0% and 70.0%. This variability highlights the diversity in voter sentiment, potentially influenced by geographic, demographic, or temporal factors.

- **Poll Count**: The higher number of polls for Trump (2,487) compared to Harris (1,253) suggests more extensive polling coverage for Trump, which may lend more stability to his average support estimate.

@@

```{r}
#| warning: false
#| message: false
#| label: fig-figure1
#| fig_caption: true
#| fig-cap:"Average support difference by state between Kamala Harris and Donald Trump. Blue indicates stronger support for Harris, red for Trump, and white for nearly equal support."
#| echo: false
# Filter data for Harris and Trump, then calculate average support by state for each
harris_data <- data %>%
  filter(candidate_name == "Kamala Harris") %>%
  group_by(state) %>%
  summarise(harris_pct = mean(pct, na.rm = TRUE))

@@

```{r}
#| warning: false
#| message: false
#| label: fig-figure2
#| fig_caption: true
#| fig-cap:"Battleground States with Narrow Polling Margins"
#| echo: false

# Calculate the average polling margin between the top two candidates in each state
important_states <- data %>%
  filter(party_binary == 1) %>%
  group_by(state, candidate_name) %>%
  summarize(avg_pct = mean(pct, na.rm = TRUE)) %>%
  arrange(state, desc(avg_pct)) %>%
  group_by(state) %>%
  mutate(margin = avg_pct - lead(avg_pct, default = 0)) %>%
  slice(1) %>%
  filter(margin < 5) # Select states with a polling margin under 5%

# Plotting the battleground states with narrow polling margins
ggplot(important_states, aes(x = reorder(state, margin), y = margin)) +
  geom_bar(stat = "identity", fill = "red") +
  coord_flip() +
  theme_minimal()
```
@fig-figure2 visualizes the polling margins between the top two candidates in various states. Each bar represents a state, with the length of the bar corresponding to the polling margin—the difference in polling percentage between the leading candidate and the runner-up. States are ordered by margin, from the narrowest at the bottom (indicating the closest contest) to the widest margin at the top. For instance, states like Arizona, Minnesota, and Michigan show extremely close polling margins, meaning that support for the top two candidates is nearly evenly split. These states are critical battlegrounds where even a small shift in voter preference could change the outcome. For example, Arizona has the smallest margin, making it a highly competitive state. Otherwise, states like Washington, California, and South Dakota have wider margins, suggesting a more clear lead for one candidate. While they are still classified as battlegrounds due to the selected margin threshold, they are less competitive than states at the bottom.

```{r}
#| label: tbl-2
#| echo: false
#| warning: false
#| message: false
#| tbl-cap: "Predicted support percentages for Kamala Harris and Donald Trump in key competitive states, showing the anticipated winner and the percentage margin of support in each state."

# Load the dataset
state_support_data <- read.csv("/Users/apple/Desktop/cleaned_president_polls.csv")

>>>>

- **Average Support**: Kamala Harris has a slight lead over Donald Trump, with an average predicted national support of 47.6% compared to Trump’s 46.85%. This narrow margin points to a competitive race, with neither candidate establishing a decisive lead across the polls.

- **Variability in Support**: The standard deviation values indicate greater variability in Trump’s support (4.41%) compared to Harris’s (4.35%). This suggests that Trump’s polling performance fluctuates more widely, which could be due to varying levels of regional support or shifts in public opinion over time.

- **Support Range**: Both candidates display a considerable range in support across polls, with Trump’s polling support varying between 27.0% and 70.0%, and Harris’s between 25.0% and 70.0%. This variability highlights the diversity in voter sentiment, potentially influenced by geographic, demographic, or temporal factors.

- **Poll Count**: The higher number of polls for Trump (1,121) compared to Harris (1,125) suggests more extensive polling coverage for Harris, which may lend more stability to her average support estimate.

@@

```{r}
#| warning: false
#| message: false
#| label: fig-figure1
#| fig_caption: true
#| fig-cap: "Average support difference by state between Kamala Harris and Donald Trump. Blue indicates stronger support for Harris, red for Trump, and white for nearly equal support."
#| echo: false

# Filter data for Harris and Trump, then calculate average support by state for each
harris_data <- data %>%
  filter(candidate_name == "Kamala Harris") %>%
  group_by(state) %>%
  summarise(harris_pct = mean(pct, na.rm = TRUE))

@@

```{r}
#| warning: false
#| message: false
#| label: fig-figure2
#| fig_caption: true
#| fig-cap: "Battleground States with Narrow Polling Margins"
#| echo: false

data <- read_csv("cleaned_president_polls.csv")
important_states <- data %>%
  group_by(state, candidate_name) %>%
  summarize(avg_pct = mean(pct, na.rm = TRUE)) %>%
  arrange(state, desc(avg_pct)) %>%
  group_by(state) %>%
  mutate(margin = avg_pct - lead(avg_pct, default = 0)) %>%
  slice(1) %>%
  filter(margin < 5) # Select states with a polling margin under 5%

# Plotting the battleground states with narrow polling margins
ggplot(important_states, aes(x = reorder(state, margin), y = margin)) +
  geom_bar(stat = "identity", fill = "red") +
  coord_flip() +
  labs(title = "Battleground States with Narrow Polling Margins",
       x = "State",
       y = "Polling Margin (%)") +
  theme_minimal()

```
@fig-figure2 visualizes the polling margins between the top two candidates in various states. Each bar represents a state, with the length of the bar corresponding to the polling margin—the difference in polling percentage between the leading candidate and the runner-up. States are ordered by margin, from the narrowest at the bottom (indicating the closest contest) to the widest margin at the top. For instance, states like Arizona, Minnesota, and Michigan show extremely close polling margins, meaning that support for the top two candidates is nearly evenly split. These states are critical battlegrounds where even a small shift in voter preference could change the outcome. For example, Arizona has the smallest margin, making it a highly competitive state. Otherwise, states like Washington, California, and South Dakota have wider margins, suggesting a more clear lead for one candidate. While they are still classified as battlegrounds due to the selected margin threshold, they are less competitive than states at the bottom.
```{r}
#| label: tbl-2
#| echo: false
#| warning: false
#| message: false
#| tbl-cap: "Predicted support percentages for Kamala Harris and Donald Trump in key competitive states, showing the anticipated winner and the percentage margin of support in each state."

# Load the dataset
state_support_data <- read_parquet("/Users/apple/Desktop/cleaned_president_polls.parquet")
